### 2014-05-23 extensions.tex

\section{Application to Monitoring Schemes}
\label{sec:extensions}

This section describes how three example monitoring schemes map to the proposed
architecture.

\subsection{Array Bounds Check}

\subsection{Uninitialized Memory Check}

\subsection{Dynamic Information Flow Tracking}

\subsection{Applicability to Other Schemes}