### 2014-05-28 tabs/filter.tex

% Operations for handling clean flags

\begin{tabular}{|l|l|l|l|}
\hline

{\bf Monitor} & {\bf Event} & {\bf Filter Condition} & {\bf Filter Operation} \\ \hline \hline

\multirow{4}{*}{BC}  
& Array allocation & Never & - \\ \cline{2-4}
& ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
& Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
& Store & Source register is clean & Mark destination memory location as clean \\
\hline\hline

\multirow{2}{*}{UMC}  
& Store & Never & - \\ \cline{2-4}
& Load & Source memory location is dirty & Do nothing \\ 
\hline\hline

\multirow{5}{*}{DIFT}  
& Software set tag & Never & - \\ \cline{2-4}
& ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
& Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
& Store & Source register is clean & Mark destination memory location as clean \\ \cline{2-4}
& Indirect jump & Source register is clean & Do nothing \\ 
\hline

\end{tabular}