### 2014-05-30 tabs/dataflow_operations.tex

% Operations for handling clean flags

\begin{tabular}{|l|l|l|l|}
\hline

{\bf Flag Type} & {\bf Event} & {\bf Filter Condition} & {\bf Dropping/Filter Operation} \\ \hline \hline

\multirow{4}{*}{Clean flag}  
& Array allocation & Never & - \\ \cline{2-4}
& ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
& Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
& Store & Source register is clean & Mark destination memory location as clean \\
\hline\hline

\multirow{4}{*}{Invalid flag}  
& Array allocation & Never & Invalidate array pointer \\ \cline{2-4}
& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
& Store & Invalid source register & Invalidate destination memory location \\
\hline%\hline

% \multirow{2}{*}{UMC}  
% & Store & Never & - \\ \cline{2-4}
% & Load & Source memory location is dirty & Do nothing \\ 
% \hline\hline
% 
% \multirow{5}{*}{DIFT}  
% & Software set tag & Never & - \\ \cline{2-4}
% & ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
% & Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
% & Store & Source register is clean & Mark destination memory location as clean \\ \cline{2-4}
% & Indirect jump & Source register is clean & Do nothing \\ 
% \hline

\end{tabular}