### 2014-05-30 tabs/monitor.tex

% Summary of monitoring operations

\begin{tabular}{|l|l|}
\hline

 {\bf Event} & {\bf Monitoring Operation} \\ \hline \hline

 Array allocation & Set base and bound metadata \\ \hline
 ALU & Copy metadata from register to register \\  \hline
 Load & Copy metadata from memory to register. Check memory access address. \\ \hline
 Store & Copy metadata from register to memory. Check memory access address. \\ 
\hline%\hline

% \multirow{2}{*}{UMC}  
%  & Store & Set metadata tag (data initialized) \\ \cline{2-3}
%  & Load & Check if metadata tag is set \\ 
% \hline\hline
% 
% \multirow{5}{*}{DIFT}  
%  & Software set tag & Set tag to a specified identifier value based on software command \\ \cline{2-3}
%  & ALU & If input tags are different, output tag is a new identifier. Otherwise, output tag is the input tag. \\ \cline{2-3}
%  & Load & Set register tag as tag of loaded data \\ \cline{2-3}
%  & Store & Set memory tag as register tag \\ \cline{2-3}
%  & Indirect jump & Check tag \\ 
%  \hline

\end{tabular}