### 2014-08-20 micro14/rebuttal.txt
Draft of MICRO 2014 rebuttal.

Review-A

-How is metadata dependence graph encoded and processed?
The full graph structure does not need to be stored. The dataflow engine
dynamically propagates the information about whether metadata is null or
invalid. Source-only dropping is enabled by only making dropping decisions on
explicit instructions which modify metadata. This does not require information
about the full dependence graph to identify.

Review-B

-Previous work has removed null-metadata (e.g., FlexiTaint).
As we mention in the paper, we are not the first work to filter out null
metadata. However, we show how a single dataflow engine can be used to provide
both this filtering mechanism as well a dropping mechanism to further reduce
overheads.

-FlexCore allows an instruction type to be not forwarded or not forwarded if
the FIFO is full.
Our implementation also filters out instruction types which are not
relevant to the monitor. Not forwarding instructions if the FIFO is full can
lead to false positives which FlexCore does not prevent.

-Baseline overheads are high. In addition, FlexCore shows much lower overheads.
One reason for our high baseline overheads is due to the need to dynamically
allocate memory space for metadata due to the large metadata we are using.
FlexCore shows much lower overheads by performing monitoring on a
reconfigurable fabric. This requires dedicated on-chip reconfigurable
fabric for this purpose. Our evaluation focuses on a core-based monitor but our
techniques could also be applied to FlexCore to further reduce its overheads.

-Information flow tracking can have false positives and negatives (e.g.,
implicit flows). Is the effect of these examined?
In this work we have focused on the false positives and negatives that arise
from dropping a portion of the information flow tracking operations. We have
not looked at the inherent false positives and negatives that arise with
specific information flow tracking schemes, such as handling implicit flows.

-Why use extra HW + extra core as opposed to a programmable HW unit?
We focused our evaluation on a core-based monitor as this has greater
flexibility and requires less hardware changes to implement.
However, our techniques could also be applied to hardware monitors such as
FlexiTaint and FlexCore to enable reduced overheads by filtering and
dropping monitoring events.

Review-C

-Why choose security monitoring instead of applications like profiling which
 can tolerate partial coverage?
Profiling would also serve as a good application for our system and we
could study such a scheme on our system in the future. We chose to focus
on security monitoring as the issues with targeting security monitoring
were a superset of those for most profiling monitors (e.g., dealing with false
positives).

Review-D

-Core idea of Section 2 (monitoring) is same as Heapmon.
Section 2 describes our baseline run-time monitoring architecture.  Our work
extends previous work on run-time monitoring, such as Heapmon and others, by
filtering out monitoring on null metadata and dropping monitoring events to
further reduce overheads.

-Claiming false negatives can be ignored undermines the soundness of the
 monitoring. Can you afford for tainted variable to be used by untrusted
 function?
Although full monitoring is always desired, if the performance overheads
of full monitoring cannot be tolerated, then false negatives are a
necessary trade off. The alternative would be to completely disable
monitoring (i.e., 100% false negatives).

-Claim that work can emulate all prior works on memory debugging and control
 flow integrity check. How can your scheme emulate every single one of them?
As the monitor is implemented in software on a core, it can support a wide
range of monitoring operations. Specifically, any monitoring technique
which performs operations in response to instruction executions should be
implementable.

Review-E

-How much additional benefit does data-flow filtering provide over
 instruction-type filtering (FlexCore)?
Like FlexCore, our system also filters out monitoring for instruction
types which are not relevant to a monitoring technique. The results we
show are the additional benefit of our technique on top of
instruction-type filtering.

-Baseline selected based on what runs on simulator instead of
 state-of-the-art (dynamic allocation due to memory issues with simulator). 
Monitoring tags can require a large memory space. This is a problem not
unique to our baseline, though it is not always addressed. We have chosen to
deal with this by performing dynamic allocation of memory for metadata rather
than statically allocating a large memory space for metadata (up to 4 GB).
This comes at a trade-off in overheads. UMC, which has smaller metadata, was
implemented by statically allocating the metadata memory and thus does not
see as high overheads as BC and DIFT. A dedicated metadata TLB could be used
to reduce these overheads.

-Processor with hardware support for information flow tracking and support
 for configurable policies can be re-purposed to provide same functionality as
 the technique proposed in the paper. 
Our technique for using a dataflow engine for filtering and dropping metadata
can be applied to both core-based monitors and dedicated hardware monitors such
as a hardware-supported information flow tracking mechanism. For a sufficiently
configurable information flow track module, it may be possible to incorporate
this into the information flow tracking policy.  However, as far as we know, we
are the first to propose using this flow tracking mechanism to provide both
filtering and dropping of monitoring events to reduce monitoring overheads. In
addition, there are monitoring schemes that may not be possible to be mapped to
information flow tracking (e.g., control flow integrity checking, hardware
error detection, etc.) in which case a separate dataflow engine, such as the
one we describe, would be needed to enable filtering and dropping.