### 2014-09-05 dropping.tex
Reorganized paper for HPCA 2015 submission.

\section{Adjustable Monitoring Infrastructure}
\label{sec:dropping}

In this section, we present an architecture that can limit the amount of
monitoring performed in order to meet a target overhead budget. We consider the
overhead budget as a soft constraint that we seek to stay below or close to,
but do not require strict guarantees that it is met. For this paper, we focus on
the performance overhead of monitoring, though other overheads could also be
considered.

% Overview of full architecture
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
    \vspace{-0.2in}
    \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
    \label{fig:dropping.overview}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:dropping.overview} shows a high-level block diagram of our
proposed architecture for enabling partial monitoring. The basic idea is to
insert hardware between the main core and the monitoring core to transparently
drop monitoring events.

There were three main challenges in designing this dropping hardware:
\begin{enumerate}
  \item \textbf{General:} The hardware needs to be applicable to a wide range of monitoring schemes.
  \item \textbf{No false positives:} False positives should never occur as a result of dropping.
  \item \textbf{Intelligent dropping:} The hardware needs to maximize the amount of useful monitoring done while staying within the overhead budget.
\end{enumerate}

In order to reduce overheads, the ``Drop'' module decides when monitoring event
should be skipped.  This decision of when to drop and which monitoring events
to dropped is discussed in more detail in Section~\ref{sec:policies}.

When an event is dropped, metadata needs to be invalidated in order to prevent
false positives. This is done using a hardware dataflow engine.
Sections~\ref{sec:dropping.false_neg_pos} and
\ref{sec:dropping.prevent_false_pos} discuss these false positive issues and
the dataflow engine in detail.

In addition, we can use the dataflow engine to filter out monitoring operations
that would be operating on invalid metadata. This maximizes the useful
monitoring that is done by the monitoring core.
Section~\ref{sec:dropping.filter} talks about how this is done using
information from the dataflow engine.

\subsection{False Negatives and False Positives}
\label{sec:dropping.false_neg_pos}

% Example of dropping
\begin{figure*}
  \begin{center}
  \subfloat[no error] {
    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_no_error.pdf}
    \label{fig:dropping.example_no_error}
  }
  \hspace{0.2in}
  \subfloat[false negative] {
    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_false_negative.pdf}
    \label{fig:dropping.example_false_negative}
  }
  \hspace{0.2in}
  \subfloat[false positive] {
    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_false_positive.pdf}
    \label{fig:dropping.example_false_positive}
  }
  \caption{Dropping}
  \label{fig:dropping.example_dropping}
  \end{center}
\end{figure*}

Our goal is to drop some monitoring operations in order to reduce the overheads
of run-time monitoring. This dropping can affect the functionality of the
monitoring schemes. There are three possible outcomes for dropping a
monitoring operation. The first is that there is no difference in operation
from the original execution. For example, if we drop line 3 from our array
bounds check example (see Figure~\ref{fig:dropping.example_no_error}), then the
check on accessing {\tt x+7} is skipped. However, this is a valid access and so
skipping the check does not change anything.

On the other hand, if the monitoring for accessing memory location {\tt y+7} on
line 4 is skipped (see Figure~\ref{fig:dropping.example_false_negative}), then
a false negative occurs. Originally, the monitor would catch this access as an
out-of-bounds access and raise an error. However, if the monitoring operation
for this is dropped, then the error is not detected. This is the trade-off that
we make in order to reduce overheads. That is, instead of either 100\% coverage
with all the associated overheads or no coverage and no overheads, we 
enable middle points of partial coverage with some fraction of the full
overheads.

The final possible outcome of dropping a monitoring operation is shown in
Figure~\ref{fig:dropping.example_dropping}. In this case, the monitoring for
line 1 is dropped, causing the bound information for pointer {\tt x} to never
be set. The result is that when the access using {\tt x} is checked on line 3,
an error will be raised. This creates a false positive where an error is
incorrectly raised. Although false negatives are part of the trade-off we make
to reduce overheads, we need to prevent false positives.

\subsection{Invalidation for Preventing False Positives}
\label{sec:dropping.prevent_false_pos}

In analyzing various monitoring schemes, we found that monitoring operations
are primarily of two types: \emph{checks} and \emph{metadata updates}. Monitors
\emph{check} certain properties to ensure correct main program execution (lines
3-4 of Figure~\ref{fig:monitoring.example_full}) and they \emph{update} metadata
for bookkeeping (lines 1-2 of Figure~\ref{fig:monitoring.example_full}). Skipping
a check operation can only cause false negatives and will never cause a false
positive (see Figure~\ref{fig:dropping.example_no_error} and
Figure~\ref{fig:dropping.example_false_negative}). Therefore, we may simply
skip a check operation. Skipping an update operation can cause false
negatives but may also cause false positives (see
Figure~\ref{fig:dropping.example_false_positive}).

% Example with invalidation
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/example_invalid.pdf}
    \caption{Invalidation flags are used to prevent false positives.}
    \label{fig:dropping.example_invalid}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata.  Thus, our approach is to mark these metadata as
invalid. Furthermore, metadata that would be derived from these dropped events
should also be marked as invalid. Figure~\ref{fig:dropping.example_invalid}
shows an example of how associating an invalid flag with metadata can be used
to avoid false positives.  In this case, when the monitoring operation to set
the metadata on {\tt x} is dropped, the metadata for {\tt x} is also marked as
invalid. The result is that when the metadata for {\tt x} is checked on line 3,
no error is raised since the metadata is invalid. Note that in this case, the
check on line 4 is also skipped. This is necessary because an error would have
been raised even if the access was in bounds.

\subsection{Dataflow Engine for Preventing False Positives}
\label{sec:dropping.arch}

% Dataflow engine high-level
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_overview.pdf}
    \caption{Hardware support for dropping.}
    \label{fig:dropping.dataflow_overview}
    \vspace{-0.1in}
  \end{center}
\end{figure}

In order to efficiently support dropping monitoring events and to prevent false
positives, we propose to insert a hardware module between the main and
monitoring core (see Figure~\ref{fig:dropping.dataflow_overview}). 
This module performs dataflow tracking on invalidation information in
order to prevent false positives. In addition, checks on invalid metadata are not useful and
thus are filtered out by the hardware module before reaching the monitor.
Similarly, updating metadata based on metadata that is invalid is also not a
productive use of the monitor. Thus, in order to save the use of the monitor
for useful operations, the hardware module also filters out update operations
based on invalid metadata.

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/dataflow_architecture.pdf}
    \vspace{-0.2in}
    \caption{Hardware architecture of the dataflow engine.}
    \label{fig:dropping.dataflow} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Figure~\ref{fig:dropping.dataflow} shows a detailed block diagram of the
hardware module inserted between the main core and the monitoring. 
Monitoring schemes typically keep metadata corresponding to the main core's
register file as well as the main core's memory space. Similarly, the dataflow
engine uses an on-chip register file to store flags for register file metadata.
In addition, a set of flags are stored that correspond to memory metadata.
These memory metadata flags are accessed through a cache and backed to main
memory. All flags are initialized as valid when the system starts.

% Address calculation and configuration
On a monitoring event, the dataflow engine reads in up to two flags in order to
determine whether the event can be filtered.  The flags to be read in typically correspond to the source
operands of the monitored instruction. For example, on a load, the dataflow
engine will typically read in the flags corresponding the memory location
accessed.
However, the architecture is designed such that the address of flags to be read
is calculated dynamically and can be configured depending on the monitoring
scheme. Specifically, there exists a configuration table that outputs a set of
control signals depending on the instruction type of the monitoring event.
These control signals are used to control a pair of address calculation units which each
contain a simple ALU. The inputs to the ALU are information from the monitored
event and a constant that is specified from the configuration table. One common
use of this address calculation is to transform addresses from byte-addressed
to word-addressed by right shifting the passed memory address by 2 bits.

\subsection{Filtering Invalid Events}
\label{sec:dropping.filter}

% Filtering decision
The decision of whether an event can be filtered out is determined using a
lookup table. This filter decision table is configured by the monitoring scheme
on system initialization. The lookup table determines whether an event can be
filtered out based on the pair of input flags read and the instruction type of
the monitored event.
For example, on an ALU instruction, if both of the source operand flags
indicate that the corresponding metadata is null, then typically the instruction can be
filtered out. Similarly, on an ALU instruction, if either of the source operand
flags indicate that their corresponding metadata is invalid, then the
instruction can also be filtered out. This filter decision is sent to the
``Filter'' block shown in Figure~\ref{fig:dropping.overview} which will simply pop
filtered entries from the FIFO and not forward them. If an entry is not
filtered, then this Filter block will pop the monitoring event entry from the
FIFO and forward it along the datapath.

% Propagating flag information
Finally, when an event is filtered out, it is necessary to propagate the flag
information for the filtered event. In addition to the decision of whether to
filter an event, the filter decision table also includes information about
whether the destination flag should be written to and what value should be
written to it. A third address calculation unit generates the write address for
this destination flag. For example, for an ALU instruction, this will
correspond to the destination register of the ALU instruction. If this ALU
instruction is filtered out due to having a pair of null source operands, then
the filter decision table will indicate that the destination register's
corresponding flag should also be marked as null.

% Flags are marked as invalid when a monitoring event is dropped due to
% insufficient overhead budget. Thus, whenever a monitoring event is dropped, the
% dropping hardware indicates this to the dataflow engine. The dataflow engine's
% write address is configured for the appropriate destination flag. When it
% receives an invalidation signal from the dropping hardware, the
% dataflow engine marks this destination flag as invalid.