### 2014-09-05 tabs/previous_overheads.tex

\begin{tabular}{|l|p{2in}|r|r|}

\hline
{\bf Name} & {\bf Description} & {\bf Average Overhead} & {\bf Worst-case Overhead} \\ \hline\hline

DIFT \cite{dift-asplos04} & Dedicated hardware for information flow tracking & 1.1\% & 23\% \\ \hline
FlexiTaint \cite{flexitaint-hpca08} & Dedicated hardware for configurable information flow tracking & 1.1\%-3.7\% & 8.7\% \\ \hline
Hardbound \cite{hardbound-asplos08} & Dedicated hardware for array bounds check & 5\%-9\% & 22\% \\ \hline
Harmoni \cite{harmoni-dsn12} & Configurable hardware for tag-based monitoring & 1\%-10\% & 20\% \\ \hline\hline

FlexCore \cite{flexcore-micro10} & FPGA-based monitor & 1.05x-1.44x & 1.84x \\ \hline
Log-Based Architecture \cite{lba-asid06} & Core-based monitor & 3.23x-7.80x & 11x \\ \hline
Log-Based Architecture Optimized \cite{lba-isca08} & Core-based monitor with hardware acceleration  & 1.02x-3.27x & 5x \\ \hline
Multi-core DIFT \cite{nagarajan-interact08} & Core-based with per-application optimized dynamic information flow tracking monitor & 1.48x & 2.2x \\ \hline\hline

LIFT \cite{lift-micro06} & Dynamic binary instrumentation for information flow tracking & 3.6x & 7.9x \\ \hline
Purfiy \cite{purify-usenix92} & Dynamic binary instrumentation for memory leaks & 2.3x & 5.5x \\ \hline
TaintCheck \cite{taintcheck-05} & Dynamic binary instrumentation for taint analysis & 10x & 27x \\ \hline

\end{tabular}