### 2014-09-12 tabs/previous_overheads.tex
Fix up citations.

%\begin{tabular}{|l|p{2in}|r|r|}
\begin{tabular}{|l|l|l|r|}

\hline
{\bf Name} & {\bf Type} & {\bf Monitoring scheme and flexibility} & {\bf Slowdown (avg./worst)} \\ \hline\hline

DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.1\% / 23\% \\ \hline
FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.1\%-3.7\% / 8.7\% \\ \hline
Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 5\%-9\% / 22\% \\ \hline
Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1\%-10\% / 20\% \\ \hline\hline

FlexCore \cite{flexcore-micro10} & Dedicated FPGA & Instruction-trace monitoring & 1.05x-1.44x / 1.84x \\ \hline
FADE \cite{fade-hpca14} & Core+Custom HW & Instruction-trace monitoring (effective when HW filters work) & 1.2x-1.8x / 3.3x \\ \hline
LBA-accelerated \cite{lba-isca08} & Multi-core+Custom HW & Instruction-trace monitoring (effective when accelerators work) & 1.02x-3.27x / 5x \\ \hline
LBA \cite{lba-asid06} & Multi-core+Custom HW & Instruction trace monitoring & 3.23x-7.80x / 11x \\ \hline \hline

Multi-core DIFT \cite{nagarajan-interact08} & SW (multithreaded) & DIFT (compiled for each application) & 1.48x / 2.2x \\ \hline

LIFT \cite{lift-micro06} & SW (DBI) & DIFT (fully flexible) & 3.6x / 7.9x \\ \hline
Purfiy \cite{purify-usenix92} & SW (DBI) & Memory leak checks (fully flexible) & 2.3x / 5.5x \\ \hline
TaintCheck \cite{taintcheck-ndsss05} & SW (DBI) & DIFT (fully flexible) & 10x / 27x \\ \hline

\end{tabular}