### 2014-10-30 hpca15/hpca2015-rebuttal.txt
Initial draft of rebuttals for HPCA 2015.

Common Questions

-Review-D,E: How does this compare to FADE?
Our main contribution in this work is an architecture which enables partial
monitoring (i.e., trade-off between coverage and overheads). We are not the
first to introduce null metadata filtering as a method for reducing monitoring
overheads. For example, FADE [11] introduces hardware to do this. However, we
show that our hardware support for partial monitoring is able to support null
metadata filtering with minor modifications. Thus, it is not necessary to
implement the full FADE hardware in order to add null metadata filtering to our
architecture for partial monitoring.

-Review-A,B,E,F: Evaluation of false negatives/error detection rate.
While we could not evaluate real errors/attacks due to difficulty in setting up
a large number of bugs and exploits and/or running real-world applications on
our simulation infrastructure, we believe that the percentage of checks
provides a good initial estimate of detection probability when errors/attacks
are uniformly distributed across checks. It will be interesting as future work
to evaluate the system with real-world bugs to see if this distribution holds
or if there exists correlations between bug locations and slack.

Review-A

-Comparison to Raksha and [Kannan DSN 2009].
Raksha and the architecture proposed in [Kannan DSN 2009]
provide hardware architectures for information flow tracking. These
architectures have very low overheads because they are designed for a specific
monitoring technique.  Although Raskha provides some programmability in the
information flow tracking policy, it does not support the wide range of
monitors supported by our architecture (e.g., race detection).

-Do you guarantee to detect violations at instruction boundaries?
We do not guarantee when violations are detected. However, previous work by
[Deng et al. ICCD 2011] has described how precise exceptions can be supported
on a monitoring architecture similar to ours.

-What is your exact definition of monitoring overhead?
Monitoring overhead is the increase in execution time of the system running
monitoring compared to without monitoring.

-Have you assessed the physical memory metadata requirements for monitors you designed?
The worst-case memory usage can be determined based on the size of metadata per
memory location. We have not evaluated the actual memory usage, but previous
work on the implemented monitoring techniques have looked at this.

-Have you considered a design where the monitor uses a RAW microprocessor like fabric 
as opposed to a FIFO for communication?
A FIFO provides a simple solution for providing buffered communication between
the main core and the monitor.

Review-B

-Design space choice: Is it possible to use another core for monitoring tasks.
The monitoring task is performed by another core (with the exception of the
FPGA-based monitor that was evaluated). However, in order to minimize the
performance impact of dropping a monitoring event, this was implemented using
custom hardware. In the past, we have explore performing these dropping
operations in software. However, this was almost as expensive as performing the
full monitoring operations and so was not effective at creating a trade-off
between coverage and performance.

Review-C

-For run time monitoring can you provide any precise guarantees regarding
 dropped requests (w/ the slack policy)?
Dropping is decided based on run-time overheads (i.e., slack). This is
difficult to predict a priori and so we are not able to give precise guarantees
on what is dropped or the coverage achieved.

-Is hardware support really necessary? Partial monitoring with software
instrumentation (Sec 5.6 or [15])? Higher overhead/coverage ratio, but with
scaling out to many users, does it matter?
A software solution is cheaper to implement. However, if the overhead/coverage
ratio is too high, then the number of users needed to detect an error could be
prohibitively high. Table 1 shows that the gap in overhead between software and
hardware solutions for full monitoring can be an order of magnitude.

-[7] has similarities to our approach.
Our baseline architecture is similar to the one proposed by [7] and other
related work. Our work shows how partial monitoring can be supported on these
architectures using a dataflow engine.

Review-D

-How much better does this do than ad-hoc filtering in terms of error?
Previous work uses filtering to reduce the overheads of monitoring. However,
they are not able to say anything about the overheads seen, especially with new
applications. The overheads can vary wildly (see Figure 9). Our architecture
allows a specified overhead to be targeted.

Review-E
See common questions.

Review-F

-Could you compare energy consumed by full and partial monitors? 
We have not evaluated the difference in energy consumed by full and partial
monitors. It is not clear which uses more energy and it largely depends on the
number of events dropped and thus the overhead target. If everything is
dropped, then less energy is used since the dataflow engine operates on
single-bit flags and is more energy-efficient than the monitor. However, if no
events are dropped or filtered out, then the dataflow engine adds energy usage
above the baseline full monitoring because it adds additional operations to be
performed.

-Tracking invalid tags through different threads in multi-processor system?
 Coherence for dataflow monitor?
Invalid tags for different threads would be tracked by the dataflow engine
through shared memory, similar to shared data for multi-threaded programs. A
multi-processor implementation of the architecture would require coherence for
the caches of the dataflow engine and monitor, similar to the coherence needed
for the cores.

-How are the sources identified for the source-only dropping scheme? How and
when do you decide drop the source?
Sources can be identified by instruction type or by the programmer/compiler.
For example, for uninitialized memory check, all store instructions are
configured to be marked as sources. Alternatively, instructions can be marked
as source events by the programmer or compiler. For example, this is done for
array bounds check when pointers are initialized to mark these points as
sources. Dropping sources can be done based on slack or using a probabilistic
approach.