### 2014-10-30 hpca15/hpca2015-rebuttal.txt
Rebuttal revision.

Common Questions

-Review-D,E: How does this compare to FADE?
Our main contribution in this work is an architecture which enables partial
monitoring (i.e., trade-off between coverage and overheads). We are not the
first to introduce null metadata filtering as a method for reducing monitoring
overheads (e.g., FADE). However, we show that our hardware support for
partial monitoring is able to support null metadata filtering with minor
modifications. Thus, it is not necessary to implement the full FADE hardware in
order to add null metadata filtering to our architecture for partial
monitoring.

-Review-A,B,E,F: Evaluation of false negatives/error detection rate.
While we could not evaluate real errors/attacks due to difficulty in setting up
a large number of bugs and exploits and/or running real-world applications on
our simulation infrastructure, we believe that the percentage of checks
provides a good initial estimate of detection probability when errors/attacks
are uniformly distributed across checks. It will be interesting as future work
to evaluate the system with real-world bugs to see if this distribution holds
or if there exists correlations between bug locations and slack.

Review-A

-Comparison to Raksha and [Kannan DSN2009].
Raksha and the architecture proposed in [Kannan DSN2009]
provide hardware architectures for information flow tracking. These
architectures have very low overheads because they are designed for a specific
monitoring technique.  Although Raskha provides some programmability in the
information flow tracking policy, it does not support the wide range of
monitors supported by our architecture (e.g., race detection).

-Do you guarantee to detect violations at instruction boundaries?
We do not guarantee when violations are detected. However, previous work by
[Deng ICCD2011] has described how precise exceptions can be supported
on a monitoring architecture similar to ours.

-Definition of monitoring overhead?
Monitoring overhead is the increase in execution time of a program running
monitoring compared to without monitoring.

-Have you assessed the physical memory metadata requirements?
The worst-case memory usage can be determined based on the size of metadata per
memory location. We have not evaluated the actual memory usage, but previous
work on the implemented monitoring techniques have looked at this.

-Have you considered using a RAW-like fabric as opposed to a FIFO for
communication?
A FIFO provided a simple solution for buffered communication between the main
core and the monitor.

Review-B

-Use another core for monitoring tasks.
The monitoring task is performed by another core. However, dropping operations
were implemented using custom hardware (i.e., dataflow engine) in order to
minimize their performance impact. We have explored performing these dropping
operations in software. However, this can be almost as expensive as performing
the full monitoring operations and so was not effective at creating a trade-off
between coverage and performance.

Review-C

-Can you provide any precise guarantees regarding dropped requests?
Dropping is decided based on run-time overheads (i.e., slack). This is
difficult to predict a priori and so we are not able to give precise guarantees
on what is dropped or the coverage achieved.

-Is hardware support really necessary? Higher overhead/coverage ratio, but with
scaling out to many users, does it matter?
Previous work that performs partial monitoring in software [15, 1] applied it
to more coarse-grained monitoring techniques instead of the fine-grained
instruction-level monitoring we target.  Table 1 shows that the gap in
overhead between software and hardware solutions for fine-grained monitoring
can be an order of magnitude.  Although a software solution is easier to
deploy, if the overhead/coverage ratio is too high, then the number of users
needed to detect an error or the minimum overhead to achieve any coverage could
be prohibitively high.

-[7] has similarities to your approach.
Our baseline architecture is similar to the one proposed by [7] and other
related work. However, our work is the first to show how partial monitoring can
be supported on these architectures using a dataflow engine.

Review-D

-How much better does this do than ad-hoc filtering in terms of error?
Previous work uses filtering to reduce the overheads of monitoring. However,
they are not able to say anything about the overheads seen, especially with new
applications. These overheads can vary wildly (see Figure 9). Our architecture
allows a specified overhead to be targeted. FlexiTaint [Venkataramani HPCA2008]
does not have a drop-on-overhead policy.

Review-E

-Should we be measuring coverage in terms of errors detected, false positives
that were generated?
In terms of errors detected, see the response in Common Questions. Our
architecture does not generate any false positives.

Review-F

-Energy consumed by full and partial monitors.
We have not evaluated the difference in energy consumed by full and partial
monitors. Dropping an event is less energy than performing full monitoring, but
the dataflow engine also adds energy overhead when not dropping. Thus, it is
not clear which uses more energy and it largely depends on the overhead
target and the number of events dropped.

-Tracking invalid tags through different threads in multi-processor system?
 Coherence for dataflow monitor?
Invalid tags for different threads would be tracked by the dataflow engine
through shared memory. A multi-processor implementation of the architecture
would require coherence for the caches of the dataflow engine and monitor,
similar to the coherence needed for the core caches.

-How are the sources identified? How do you decide to drop the source?
Sources are identified by instruction type.
For example, for uninitialized memory check, all store instructions are
configured to be marked as sources. In addition, custom instructions used for
setting metadata (e.g., setting array bounds metadata) are always marked as
sources. Dropping sources can be done based on slack or using a probabilistic
approach.