### 2015-01-05 policies.tex
Discussion on sub-flow dropping.

\section{Dropping Policies} \label{sec:policies}

In order to use partial monitoring to enable adjustable overheads, we must also
specify a policy for when and which monitoring events are dropped. In this
section, we discuss some of the options and trade-offs for dropping policies.
We split this decision into two components: \begin{enumerate} \item When do we
need to drop events in order to enable reduced overheads?
(Section~\ref{sec:policies.when}) \item Which events should be dropped?
(Section~\ref{sec:policies.which}) \end{enumerate}

% Our goal with enabling partial monitoring is to allow for a specified
% overhead target or budget to be met while maximizing the amount of monitoring
% performed.  This is done by dynamically dropping a portion of the monitoring
% operations. In this section, we discuss in detail how this dropping decision
% is made. Specifically, we first discuss how to decide when dropping should be
% performed Section~\ref{sec:policies.slack}. Next, we investigate trade-offs
% created by selecting which events are dropped
% (Section~\ref{sec:policies.events}). Different policies in determining which
% events can be dropped create a trade-off between how closely the overhead
% budget is met and the monitoring coverage achieved.

\subsection{Deciding When to Drop} \label{sec:policies.when}

% Slack
\begin{figure} \begin{center}
\includegraphics[width=\columnwidth]{figs/slack.pdf} \vspace{-0.2in}
\caption{Slack and its effect on monitoring over time.}
\label{fig:policies.slack} \vspace{-0.1in} \end{center} \end{figure}

In this section, we discuss two possible ways to determine when events should
be dropped.  The first possibility is to probabilistically drop events. By
setting the probability of dropping events appropriately, overheads can be
reduced. This works well for enabling partial monitoring for cooperative
testing and debugging since the randomness allows different users and runs to
monitor different portions of the program. However, using a probabilistic
dropping policy can make it difficult to meet a target overhead without prior
profiling.

Alternatively, we can specify a target overhead and estimate, at run-time, the
overheads of monitoring in order to decide whether dropping is needed.  The
overhead budget is specified as a percentage of the main program's execution
cycles without monitoring. In order to estimate overheads at run-time, we
define \emph{slack} as the number of cycles of monitoring overhead that can be
incurred while staying within the budget target. Slack is essentially the
difference between the actual overhead seen and the budget specified. Slack is
generated as the main program runs and consumed as monitoring overheads occur.
For example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles.  In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:policies.slack} shows an example of how slack can change over
time.  In this slack-based policy, if the slack falls below zero (i.e., the
overhead budget is exceeded), then monitoring events are dropped.

% % Slack tracking module \begin{figure} \begin{center}
% \includegraphics[width=\columnwidth]{figs/stm.pdf} \vspace{-0.2in}
% \caption{Slack tracking and drop decision hardware.} \label{fig:policies.stm}
% \vspace{-0.1in} \end{center} \end{figure}

% Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
% keeping track of slack. 
Slack can be easily measured in hardware by using a counter that increments on
every $k$-th cycle of the main core (e.g., every 5th cycle for a 20\% target
budget). 
% This $k$ can be calculated by taking the reciprocal of the target budget. For
% example, if the target budget is 20\%, then the counter increments on every
% 5th clock cycle. 
The value of this counter is the accumulated slack. Whenever the main core is
stalled due to the monitor, the measured slack is decremented. It is difficult
to precisely determine the entire impact of monitoring on the main core due to
the difficulty in measuring certain overheads such as contention for shared
memory.  However, we have found that using only the stalls due to FIFO back
pressure works well in practice.

\subsection{Deciding Which Events to Drop} \label{sec:policies.which}

% % Varying slack's impact on coverage for UMC \begin{figure} \begin{center}
% \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf} \vspace{-0.3in}
% \caption{Example dependence graph for metadata. Square nodes represent events
% where checks are performed.} \label{fig:policies.dataflow_graph}
% \vspace{-0.1in} \end{center} \end{figure}

\begin{figure} \begin{center} \subfloat[Unrestricted Dropping]{
\includegraphics[width=\columnwidth]{figs/unrestricted_drop.pdf}
\label{fig:policies.all_drop} } \vspace{-0.1in} \\ \subfloat[Source-Only
Dropping]{ \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
\label{fig:policies.source_drop} } \vspace{-0.1in} \\ \subfloat[Sub-Flow
Dropping]{ \includegraphics[width=\columnwidth]{figs/subflow_drop.pdf}
\label{fig:policies.subflow_drop} } \end{center} \caption{Comparison of
dropping policies using metadata dependence graphs.  Square nodes represent
events where check are performed. Blue (dark) nodes indicate which nodes can be
dropped.} \label{fig:policies.policies} \end{figure}

In addition to deciding when dropping is required, trade-offs also exist in
deciding which events should be dropped.  The simplest policy is to drop any
monitoring events when slack is less than or equal to zero.  However, this can
result in wasted work. For example, consider the metadata dependence graph
shown in Figure~\ref{fig:policies.all_drop}. Here, an edge from node {\tt A} to
node {\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be dropped. Square nodes
indicate events where monitoring checks are performed. In the example, suppose
that event {\tt E} is meant to perform a check operation but is dropped.  In
this situation, the monitoring operations that were done for events {\tt C} and
{\tt D} were wasted since their results were not used for any monitoring
checks.  That is, by the time we decide to drop event {\tt E}, we have already
updated metadata for events {\tt C} and {\tt D} even though they are no longer
needed.

An alternative dropping policy which eliminates this wasted work is to only
make dropping decisions at the root or source of these metadata flows. That is,
we will decide to either monitor or not monitor an entire metadata flow. An
example of these source nodes is the monitoring done to initialize base and
bounds information on {\tt malloc} for an array bounds check. These source
nodes are typically easy to identify by the dropping hardware because they
correspond to special instructions that are used to set up metadata
information. Thus, it is not necessary to generate and analyze the monitoring
dependence graph to identify source nodes. We refer to this dropping decision
policy as \emph{source-only
dropping} and we refer to the previous policy of dropping any event as
\emph{unrestricted dropping}.

Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage at a given overhead. However, because of the coarser-grained decision, it
may be more difficult to closely match overhead targets. We expect source-only dropping 
to work well when there are a large number of independent metadata flows.
In addition, using a probabilistic dropping policy works poorly with
unrestricted dropping. Since every event in a dependent chain (e.g., events {\tt
A} through {\tt E}) needs to be monitored in order for the monitoring check to
be useful, randomly dropping any event is likely to cause the final check to be
invalid by dropping at least one event in the chain.

More complex dropping policies can be implemented given more detailed
information about the monitoring dependence graph. Here, we describe one such
policy, which we call \emph{sub-flow dropping}. Sub-flow dropping makes
dropping decisions at a granularity in between unrestricted dropping and
source-only dropping (see Figure~\ref{fig:policies.subflow_drop}). Sub-flow
dropping allows source nodes to be dropped and nodes after branch points in the
monitoring dependence graph to be dropped. From
Figure~\ref{fig:policies.subflow_drop}, this corresponds to nodes {\tt A}, {\tt
C}, and {\tt F}. The basic intuition is that dropping certain portions of the
monitoring dependence graph still result in no wasted work. For example,
dropping node {\tt C} causes nodes {\tt D} and {\tt E} to be skipped. However,
performing monitoring on the remaining nodes allows the check at node {\tt G}
to be performed with no wasted work. Similarly, dropping {\tt F} allows node
{\tt E} to be checked with no wasted work.

Depending on the target application of partial monitoring, different policies
are more applicable.
For applications where closely matching an overhead target is important, a
slack-based, unrestricted dropping policy is appropriate. However, if matching
the overhead target is not as important, then a slack-based, source-only dropping
policy could provide better coverage. 
Finally, if the goal is to use partial monitoring to enable cooperative
debugging and testing with very low overheads, then a probabilistic,
source-only dropping policy can be used to provide good total coverage over
multiple runs.
Sub-flow dropping enables a design point in between unrestricted dropping and
source-only dropping. Sub-flow dropping matches overheads better than
source-only dropping and provides better coverage than unrestricted dropping.
For a probabilistic policy, sub-flow dropping allows better overhead matching
while still allowing good total coverage over multiple runs, though at a slower
rate than source-only dropping.

% Trade-off between source-drop and unrestricted dropping
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/policy_trade_off.pdf}
%     \vspace{-0.3in}
%     \caption{Slack tracking and drop decision hardware.}
%     \label{fig:policies.trade_off}
%     \vspace{-0.2in}
%   \end{center}
% \end{figure}
% 
% Another possible method is to make the dropping decisions somewhere in between
% source-only dropping and unrestricted dropping. From the graph in 
% Figure~\ref{fig:policies.dataflow_graph}, we can see that if the dropping
% decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
% work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
% monitoring on that flow. Similarly, event {\tt F} creates a decision point for
% the bottom flow that will result in no wasted work and dropping event {\tt A}
% allows the entire shown metadata flow to be dropped with no wasted work. If the
% program can be
% analyzed or profiled to identify these instructions which, if dropped, lead to
% no wasted work, then this information can be used at run-time to minimize the
% amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
% Figure~\ref{fig:policies.min_drop} shows the points were dropping decisions are
% made for this policy.
% Sub-flow dropping is a coarser-grained decision than unrestricted dropping, but
% finer-grained that source-only dropping. Thus, its ability to match overheads is
% expected to sit between these other two policies. Similarly, there still exists
% edge cases where wasted work can occur, but typically we expect much less
% wasted work than unrestricted 
% dropping. Thus, we expect the coverage achieved by sub-flow dropping for a
% certain overhead to fall between source-only and unrestricted dropping. These three
% dropping policies create a trade-off space between coverage achieved and
% ability to match an overhead target.
% Figure~\ref{fig:policies.trade_off} summarizes this trade-off between coverage
% and ability to match overhead.